---
title: "compare_EC_MGS_PIC"
auteur: "Dinh-Phuc Hoang"
date: "2025-03-05"
---

```{r}
library(tidyverse)
library(dplyr)
library(phyloseq)
library(VennDiagram)
library(reshape2)
library(ggplot2)
```

```{r}
EC_P2 = read.delim("Data FONCTIOMIC/EC_P2.txt", sep ="\t")
EC_MGS = read.delim("Data FONCTIOMIC/EC_MGS.txt", sep ="\t")
EC_P2 = as.matrix(EC_P2)
EC_MGS = as.matrix(EC_MGS)
EC_number_class = read.delim("description_mapfiles/EC_number_class.txt", sep ="\t")
rownames(EC_number_class) = EC_number_class$EC_number
EC_class = as.data.frame(EC_number_class[,-1])
colnames(EC_class) = "EC_class"
rownames(EC_class) = EC_number_class$EC_number
```
Correlation of Spearman on EC abundance between PICRUSt2 and MGS (Robust)
```{r}
corr = cor.test(EC_MGS, EC_P2, method = "spearman", exact = FALSE)
cor.1 = cor(t(EC_MGS), t(EC_P2), method = "spearman")
diag_cor = diag(cor.1)
diag_cor = as.data.frame(diag_cor)
names(diag_cor) = "Spearman_correlation"
diag_cor = rownames_to_column(diag_cor, var = "EC_number")
diag_cor.1 = merge(diag_cor, EC_number_class, by = "EC_number", join = "left")
```
Boxplot for Spearman correlation coeffectient between EC abundance from PICRUSt2 and MGS
```{r}
ggplot(diag_cor.1, aes(x = Spearman_correlation, y = EC_class,group = EC_class)) +
  geom_boxplot(fill="gray", alpha=0.2)+
  ggtitle("Spearman correlation between EC abundance from PICRUSt2 and MGS") +
  xlab("Spearman correlation") +
  ylab("EC class")
```
Venn Diagramm for EC detected between PICRUSt2 and MGS
```{r}
venn.diagram(
        x = list(rownames(tb_ec_cleaned), rownames(p2_EC)),
        category.names = c("MGS" , "PICRUSt2"),
        filename = '#14_venn_diagramm.png',
        output=TRUE,
        imagetype="png" ,
        lwd = 2,
        lty = 'blank',
        fill = c("cornflowerblue", "pink"),
        cat.fontface = "bold",
        cat.default.pos = "outer",
        cat.fontfamily = "sans",
)
```
Import ALDEx2 for EC abundance (PICRUSt2 + MGS)
```{r}
EC_P2_P_value = read.delim("ALDEx2/p2_EC_aldex.txt", sep ="\t")
EC_MGS_P_value = read.delim("ALDEx2/MGS_EC_aldex.txt", sep ="\t")
EC_number_class = read.delim("description_mapfiles/EC_number_class.txt", sep ="\t")
```
Correlation of Spearman on P value between PICRUSt2 and MGS (Robust)
```{r}
c = intersect(rownames(EC_P2_P_value), rownames(EC_MGS_P_value))
EC_P2_P_value = EC_P2_P_value[c,1:4]
EC_MGS_P_value = EC_MGS_P_value[c,1:4]
EC_P2_P_value = as.matrix(EC_P2_P_value)
EC_MGS_P_value = as.matrix(EC_MGS_P_value)
cor.2 = cor(t(EC_MGS_P_value), t(EC_P2_P_value), method = "spearman")
corr.2 = cor.test(EC_MGS_P_value, EC_P2_P_value, method = "spearman", exact = FALSE)
diag_cor.2 = diag(cor.2)
diag_cor.2 = as.data.frame(diag_cor.2)
diag_cor.2$EC_number = rownames(cor.2)
class_cor.2 = merge(diag_cor.2, EC_number_class, by = "EC_number")
```
Plot for Spearman correlation of P value between EC differential abundance from PICRUSt2 and MGS
```{r}
ggplot(class_cor.2, aes(x = class_cor.2$diag_cor.2, y = EC_class, group = EC_class)) +
  geom_boxplot(fill="gray", alpha=0.2)+
  ggtitle("Spearman correlation of P value between EC differential abundance from PICRUSt2 and MGS") +
  xlab("Spearman correlation") +
  ylab("EC class")
```
Correlation des différences (Control contre Residues) entre PICRUSt2 et MGS
```{r}
#Au temps 0
plot(dif_0[overlapping_EC,], P2_dif_0[overlapping_EC,],
     xlab = "Control vs Residues from MGS",
     ylab = "Control vs Residues from PICRUSt2",
     main = "Correlation of differences EC abundance between Control vs Residues from PICRUSt2 and MGS at T0")
#testé avec qqnorm pour voir si les données sont normalement distribuées (la réponse est non) => corrélation de Spearman
cor_0 = cor(t(dif_0[overlapping_EC,]),t(P2_dif_0[overlapping_EC,]), method = "spearman")
corr_0 = cor.test(t(dif_0[overlapping_EC,]),t(P2_dif_0[overlapping_EC,]), method = "spearman")
diag_cor_0 = diag(cor_0)
diag_cor_0 = as.data.frame(diag_cor_0)
diag_cor_0$EC_number = rownames(diag_cor_0)
diag_cor_0 = merge(diag_cor_0, t, by = "EC_number")
ggplot(diag_cor_0, aes(x = diag_cor_0, y = EC_class, group = EC_class)) +
  geom_boxplot(fill="gray", alpha=0.2)+
  ggtitle("Spearman correlation of differences EC abundance between Control vs Residues from PICRUSt2 and MGS at T0") +
  xlab("Spearman correlation") +
  ylab("EC class")
#Au temps 51
plot(dif_51[overlapping_EC,], P2_dif_51[overlapping_EC,],
     xlab = "Control vs Residues from MGS",
     ylab = "Control vs Residues from PICRUSt2",
     main = "Correlation of differences EC abundance between Control vs Residues from PICRUSt2 and MGS at T51")
cor_51 = cor(t(dif_51[overlapping_EC,]),t(P2_dif_51[overlapping_EC,]), method = "spearman")
corr_51 = cor.test(t(dif_51[overlapping_EC,]),t(P2_dif_51[overlapping_EC,]), method = "spearman")
diag_cor_51 = diag(cor_51)
diag_cor_51 = as.data.frame(diag_cor_51)
diag_cor_51$EC_number = rownames(diag_cor_51)
diag_cor_51 = merge(diag_cor_51, t, by = "EC_number")
ggplot(diag_cor_51, aes(x = diag_cor_51, y = EC_class, group = EC_class)) +
  geom_boxplot(fill="gray", alpha=0.2)+
  ggtitle("Spearman correlation of differences EC abundance between Control vs Residues from PICRUSt2 and MGS at T51") +
  xlab("Spearman correlation") +
  ylab("EC class")
```

```{r}
top_EC_P2=read.delim("top_EC_P2.txt", sep ="\t")
top_EC_P2 = as.data.frame(top_EC_P2[,1])
colnames(top_EC_P2) = "log2Fold change"
top_EC_MGS=read.delim("top_EC_MGS.txt", sep ="\t")
pheatmap(top_EC_P2,
show_rownames = FALSE,
show_colnames = FALSE,
annotate_row = EC_class,
cluster_rows = TRUE,
cluster_cols = FALSE,
main = "Top ECs having a significants change from PICRUSt2")
```