---
title: "FONCTIOMIC"
auteur: "Dinh-Phuc Hoang"
date: "2025-02-25"
---


```{r}
library(tidyverse)
library(dplyr)
library(phyloseq)
library(ALDEx2)
library(DESeq2)
library(pheatmap)
library(BiocParallel)
library(Hotelling)
register(SnowParam(4))
```

```{r}
tb_ec_raw = read_tsv("Data FONCTIOMIC/counts_by_EC_number.tsv")
tb_ec = read_tsv("Data FONCTIOMIC/counts_by_EC_number.norm.tsv")
tb_gene = read_tsv("Data FONCTIOMIC/TOY_ALL.counts.norm_cleaned.tsv")
tb_cog = read_tsv("Data FONCTIOMIC/TOY_PROKKA_02252022_all3000bp.tsv")
meta = read_tsv("Data FONCTIOMIC/complete_metadata_table.tsv")
meta = as.data.frame(meta)
row.names(meta) = meta$sample_id
EC_number_class = read.table("description_mapfiles/EC_number_class.txt", header = TRUE, sep = "\t")
```
Tidy up EC table
```{r}
tb_ec = as.data.frame(tb_ec)
row.names(tb_ec) = paste0("EC:", tb_ec$EC_number)
tb_ec_cleaned = tb_ec %>% filter(!str_detect(rownames(tb_ec), "-$"))
tb_ec_cleaned = tb_ec_cleaned[,-c(1,50)]
tb_ec_cleaned = as.matrix(tb_ec_cleaned)
```
PCA for EC abundance

```{r}
tb_ec_cleaned_pca = t(tb_ec_cleaned)
ec_pca = vegan::rda(tb_ec_cleaned_pca)
summary(ec_pca)
```
PCA plot for EC abundance
```{r}
ec_pca_plot = as.data.frame(ec_pca$CA$u[,1:2])
ec_pca_plot = cbind(ec_pca_plot, meta$Category)
colnames(ec_pca_plot) <- c("PC1","PC2","Groups")
ggplot(ec_pca_plot, aes(x = PC1, y = PC2, color = Groups, shape = Groups)) +
  ggtitle("PCA of EC abundance") +
  geom_point(alpha=1) +
  xlab("PC1: 9%")+
  ylab("PC2: 1%") +
  stat_ellipse(geom="polygon",type ="norm", level=0.9, alpha=0.1,
  aes(fill = Groups),
  color =NA) +
theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
```
overlap EC between MSG and PICRUsT2 on FONCTIOMIC
```{r}
overlapping_EC = intersect(rownames(p2_EC), rownames(tb_ec_cleaned))
overlapping_sample = intersect(colnames(p2_EC), colnames(tb_ec_cleaned))
EC_P2 = p2_EC[overlapping_EC, overlapping_sample]
write.table(EC_P2, "Data FONCTIOMIC/EC_P2.txt",sep="\t")
EC_MGS = tb_ec_cleaned[overlapping_EC, overlapping_sample]
write.table(EC_MGS, "Data FONCTIOMIC/EC_MGS.txt",sep="\t")
```
Statistical analysis with ALDEx2 between 4 groups
```{r}
meta$Category <- as.factor(meta$Category)
meta$Category <- relevel(meta$Category, ref = "Control Cropland")
mm = model.matrix(~ meta$Category + meta$time, data = meta)
tb_ec_cleaned = as.matrix(tb_ec_cleaned)
tb_ec_cleaned = round(tb_ec_cleaned,0)
x = aldex.clr (tb_ec_cleaned, mm, mc.samples = 128, denom = "all",useMC=TRUE)
glm.test = aldex.glm(x, mm)
glm.effect <- aldex.glm.effect(x,useMC = TRUE)
MGS_EC_aldex = glm.test[,13:20]
write.table(MGS_EC_aldex, "ALDEx2/MGS_EC_aldex.txt", sep = "\t", quote = FALSE, row.names = TRUE, col.names = TRUE)
```
**bold**#Statistical analysis with DESeq2 between 4 groups
Slice the data by time #time = 0
```{r}
tb_ec_cleaned = round(tb_ec_cleaned,0)
meta$Category = as.factor(meta$Category)
meta$Category = relevel(meta$Category, ref = "Control Cropland")
meta$time = as.factor(meta$time)
meta$wheat_straw_amendment = as.factor(meta$wheat_straw_amendment)
meta$wheat_straw_amendment = relevel(meta$wheat_straw_amendment, ref = "Control")
dds_MGS = DESeqDataSetFromMatrix(countData =tb_ec_cleaned,
                                  colData = meta,
                                  design = ~ wheat_straw_amendment + time, tidy = FALSE)
DAA_EC_MGS = DESeq(dds_MGS, parallel = TRUE)
dds_MGS = vst(dds_MGS,blind = FALSE)
```
Extracting results
```{r}
res_EC = results(DAA_EC_MGS, parallel = TRUE, contrast = c("wheat_straw_amendment",
"Control",
"Amended"))
head(res_EC)
summary(res_EC)
resultsNames(DAA_EC_MGS)
resSig_EC = subset(res_EC, padj < 0.05 & pvalue < 0.05)
head(resSig_EC[ order(resSig_EC$log2FoldChange, decreasing = TRUE), ])
dds_MGS_vst = assay(dds_MGS)
dds_MGS_vst = as.data.frame(dds_MGS_vst)
dds_MGS_vst$EC_number = rownames(dds_MGS_vst)
dds_MGS_vst = subset(dds_MGS_vst, EC_number %in% rownames(resSig_EC))
```

```{r}
pcaData = plotPCA(dds_MGS, intgroup = c("Category", "time"), returnData = TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category, shape = time)) +
  geom_point(size =3) +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of MGS") +
 scale_color_manual(values = c("#FF6F91",
                                "#C34A36",
                                "#2C73D2",
                                "#161a19"))+
  scale_shape_manual(values = c(16,15,17,18))+
  scale_fill_discrete(breaks = c("Control Cropland","Amended Cropland","Control Grassland","Amended Grassland"))+
  guides(color = guide_legend(title = "Groups"),shape = guide_legend(title = "Time")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
```
Sample distances
```{r}
dds_MGS_vst = as.matrix(dds_MGS_vst[,-c(49)])
dds_MGS_vst = dds_MGS_vst[order(rownames(dds_MGS_vst)),
order(colnames(dds_MGS_vst))]
mat = dds_MGS_vst
mat = mat - rowMeans(mat)
dists_row = vegan::vegdist((dds_MGS_vst), method = "bray")
dists_col = vegan::vegdist(t(dds_MGS_vst), method = "bray")
```
Heatmap of the expression of ECs across the groups (by time and land types)
```{r}
anno_col = as.data.frame(meta[,c("land_use","wheat_straw_amendment","time")])
colnames(anno_col) = c("Land types","Treatments","Time")
anno_col$Time = as.factor(anno_col$Time)
anno_row = t[,2]
anno_row = as.data.frame(anno_row)
colnames(anno_row) = "EC Class"
rownames(anno_row) = t$EC_number
pheatmap(mat, clustering_distance_rows = dists_row, 
              clustering_distance_cols = dists_col,
              show_rownames = FALSE, 
              cutree_rows = 5, #9, #5
              cutree_cols = 3,
              annotation_col = anno_col, 
              annotation_row = anno_row)
```
Define of variables for subset matrix of EC
```{r}
for (groups in c("Control Cropland","Amended Cropland","Control Grassland","Amended Grassland")){
  for (time in c("0","3","51","125")){
  assign(paste0(gsub(" ","_",groups),"_",time), str_sort(rownames(meta[meta$Category == groups & meta$time == time,])))
  }
}
crop_0 = c(Control_Cropland_0,Amended_Cropland_0)
crop_3 = c(Control_Cropland_3,Amended_Cropland_3)
crop_51 = c(Control_Cropland_51,Amended_Cropland_51)
crop_125 = c(Control_Cropland_125,Amended_Cropland_125)
grass_0 = c(Control_Grassland_0,Amended_Grassland_0)
grass_3 = c(Control_Grassland_3,Amended_Grassland_3)
grass_51 = c(Control_Grassland_51,Amended_Grassland_51)
grass_125 = c(Control_Grassland_125,Amended_Grassland_125)
```
Subset matrix of EC en fonction of time and traitement
```{r}
for (groups in c("Control Cropland","Amended Cropland","Control Grassland","Amended Grassland")) {
  for (time in c("0","3","51","125")) {
    assign(paste0("mat_", gsub(" ", "_", groups), "_", time), tb_ec_cleaned[, get(paste0(gsub(" ", "_", groups), "_", time))])
  }
}
```
Calcul de différence entre les matrices
```{r}
#Cropland
dif_C_0 = mat_Control_Cropland_0 - mat_Amended_Cropland_0
dif_C_3 = mat_Control_Cropland_3 - mat_Amended_Cropland_3
dif_C_51 = mat_Control_Cropland_51 - mat_Amended_Cropland_51
dif_C_125 = mat_Control_Cropland_125 - mat_Amended_Cropland_125
#Grassland
dif_G_0 = mat_Control_Grassland_0 - mat_Amended_Grassland_0
dif_G_3 = mat_Control_Grassland_3 - mat_Amended_Grassland_3
dif_G_51 = mat_Control_Grassland_51 - mat_Amended_Grassland_51
dif_G_125 = mat_Control_Grassland_125 - mat_Amended_Grassland_125
```

```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        EC_number = character())
tb_ec_cleaned = round(tb_ec_cleaned,0)
tb_ec_cleaned = as.data.frame(tb_ec_cleaned)
for ( dataset in c("crop_0", "crop_3", "crop_51", "crop_125")){
  counts_matrix <- as.matrix(tb_ec_cleaned[,get(dataset)],rownames = TRUE)
  sample_data <- meta[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ wheat_straw_amendment, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$EC_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.1 & pvalue < 0.05)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, EC_number_class, by = "EC_number")
res4_df[23,9] = "Hydrolases" 
res4_df[9,9] = "Translocases" 
res4_df[11,9] = "Oxireductases" 
res4_df[33,9] = "Transferases"
tb_ec_crops = tb_ec_cleaned[res4_df$EC_number,c(crop_0,crop_3,crop_51,crop_125)]
```
```{r}
dist_a_vs_c = vegan::vegdist(t(tb_ec_cleaned), method = "bray")
tree_a_vs_c = hclust(dist_a_vs_c, method = "complete")
cutree_a_vs_c = cutree(tree_ec, k = 3 )
table(cutree_a_vs_c)
```
```{r}
dist_ec = vegan::vegdist(tb_ec_clr, method = "bray")
tree_ec = hclust(dist_ec, method = "complete")
#cut into arbitrary clusters
cutree_ec = cutree(tree_ec, k = 7 )
table(cutree_ec)
```
```{r}
tb_ec_clr = clr(tb_ec_cleaned)
anno_col = as.data.frame(meta[,c("Category","time")])
colnames(anno_col) = c("Groups","Time")
rownames(anno_col) = rownames(meta)
anno_col$Groups = as.factor(anno_col$Groups)
anno_col$Time = as.factor(anno_col$Time)
vect_colours = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd", 
                "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2")
anno_col = cutree_a_vs_c
anno_row = res4_df[,9]
anno_row = as.data.frame(anno_row)
anno.row = cutree_ec
colnames(anno_row) = "EC Class"
rownames(anno_row) = rownames(tb_ec_cleaned)
pheatmap(tb_ec_crops, 
              show_rownames = FALSE, 
              cutree_rows = 7, #9, #5
              cutree_cols = 3,
              annotation_colors = list(Groups = vect_colours),
              annotation_col = anno_col, 
              annotation_row = anno_row)
```
#Rerun chuck above for reset tb_ec_cleaned to intial
```{r}
tb_ec_cleaned = round(tb_ec_cleaned,0)
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        EC_number = character())
tb_ec_cleaned = as.data.frame(tb_ec_cleaned)
for ( dataset in c("grass_0", "grass_3", "grass_51", "grass_125")){
  counts_matrix <- as.matrix(tb_ec_cleaned[,get(dataset)],rownames = TRUE)
  sample_data <- meta[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ wheat_straw_amendment, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$EC_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.1 & pvalue < 0.05)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, EC_number_class, by = "EC_number")

```