---
title: "FONCTIOMIC_MGS"
auteur: "Dinh-Phuc Hoang"
date: "2025-03-21"
output:
  html_document:
    code_folding: hide
---

#Loading packages
```{r}
library(tidyverse)
library(dplyr)
library(phyloseq)
library(DESeq2)
library(ggplot2)
library(ComplexHeatmap)
library(gridExtra)
library(GGally)
library(ggpubr)
library(BiocParallel)
library(FactoMineR)
library(microbiomeMarker)
library(ade4)
register(SnowParam(4))
```
#Data loading
```{r}
tb_ec = read_tsv("Data FONCTIOMIC/counts_by_EC_number.norm.tsv")
meta = read_tsv("Data FONCTIOMIC/complete_metadata_table.tsv")
meta = as.data.frame(meta)
row.names(meta) = meta$sample_id
meta$wheat_straw_amendment = factor(meta$wheat_straw_amendment, levels =  c("Control","Amended"))
EC_number_class = read.table("description_mapfiles/EC_number_class.txt", header = TRUE, sep = "\t")
rownames(EC_number_class) = EC_number_class$EC_number
tb_cog = read_tsv("Data FONCTIOMIC/counts_by_COG.raw.tsv")
COG_anno = read_rds("description_mapfiles/COG_anno.rds")
COG_anno$COG_number = rownames(COG_anno)
tb_pathways = read_delim("Data FONCTIOMIC/pathways_mgs.tabular", delim = "\t")
```
##Define of variables for subset matrix
```{r}
for (groups in c("Control Cropland","Amended Cropland","Control Grassland","Amended Grassland")){
  for (time in c("0","3","51","125")){
  assign(paste0(gsub(" ","_",groups),"_",time), str_sort(rownames(meta[meta$Category == groups & meta$time == time,])))
  }
}
list.replicate = list("CC_0" = Control_Cropland_0,
                      "CC_3" =Control_Cropland_3,
                      "CC_51" =Control_Cropland_51,
                      "CC_125" =Control_Cropland_125,
                      "CR_0" =Amended_Cropland_0,
                      "CR_3" =Amended_Cropland_3,
                      "CR_51" =Amended_Cropland_51,
                      "CR_125" =Amended_Cropland_125,
                      "PC_0" =Control_Grassland_0,
                      "PC_3" =Control_Grassland_3,
                      "PC_51" =Control_Grassland_51,
                      "PC_125" =Control_Grassland_125,
                      "PR_0" =Amended_Grassland_0,
                      "PR_3" =Amended_Grassland_3,
                      "PR_51" =Amended_Grassland_51,
                      "PR_125" =Amended_Grassland_125)
crop_0 = c(Control_Cropland_0,Amended_Cropland_0)
crop_3 = c(Control_Cropland_3,Amended_Cropland_3)
crop_51 = c(Control_Cropland_51,Amended_Cropland_51)
crop_125 = c(Control_Cropland_125,Amended_Cropland_125)
grass_0 = c(Control_Grassland_0,Amended_Grassland_0)
grass_3 = c(Control_Grassland_3,Amended_Grassland_3)
grass_51 = c(Control_Grassland_51,Amended_Grassland_51)
grass_125 = c(Control_Grassland_125,Amended_Grassland_125)
control_0 = c(Control_Cropland_0,Control_Grassland_0)
control_3 = c(Control_Cropland_3,Control_Grassland_3)
control_51 = c(Control_Cropland_51,Control_Grassland_51)
control_125 = c(Control_Cropland_125,Control_Grassland_125)
amended_0 = c(Amended_Cropland_0,Amended_Grassland_0)
amended_3 = c(Amended_Cropland_3,Amended_Grassland_3)
amended_51 = c(Amended_Cropland_51,Amended_Grassland_51)
amended_125 = c(Amended_Cropland_125,Amended_Grassland_125)
```

#**Data description**
* tb_ec: abundance of EC number in the output of MGS
* meta: metadata of samples
* EC_number_class: description of EC number
* tb_ec_cleaned: abundance of EC number in the output of MGS after cleaning
#Part I : _Analysis of EC abundance in the output of MGS_
##Tidy up the table abundance of EC
```{r}
tb_ec = as.data.frame(tb_ec)
row.names(tb_ec) = paste0("EC:", tb_ec$EC_number)
tb_ec_cleaned = tb_ec %>% filter(!str_detect(rownames(tb_ec), "-$"))
tb_ec_cleaned = tb_ec_cleaned[,-c(1,50)]
#overlapping_ec = read_rds("Data FONCTIOMIC/overlapping_ec.rds")
#tb_ec_cleaned = tb_ec_cleaned[overlapping_ec,]
```
##PCA analysis (VST)
```{r}
meta$Category = as.factor(meta$Category)
meta$Category = relevel(meta$Category, ref = "Control Cropland")
meta$time = as.factor(meta$time)
meta$time = relevel(meta$time, ref = "0")
meta$wheat_straw_amendment = as.factor(meta$wheat_straw_amendment)
meta$wheat_straw_amendment = relevel(meta$wheat_straw_amendment, ref = "Control")
tb_ec_vst = round(tb_ec_cleaned,0)
dds_MGS = DESeqDataSetFromMatrix(countData = tb_ec_vst,
                                  colData = meta,
                                  design = ~ Category + time, tidy = FALSE)
vst_MGS = vst(dds_MGS,blind = FALSE) # variance stabilizing transformation
#PCA plot
pcaData = plotPCA(vst_MGS, intgroup = c("Category", "time"), returnData = TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of MGS (VST)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_EC_MGS.png", width = 8, height = 10, dpi = 300)
```
##PCA analysis (Raw data)
```{r}
res.pca = PCA(tb_ec_cleaned, scale.unit = FALSE, graph = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(res.pca$eig[1:2,2], 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of MGS (Data raw)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_EC_MGS_raw.png", width = 8, height = 10, dpi = 300)
```
##PCA analysis (CLR transformation of package microbiomeMarker)
```{r}
tb_ec_clr = microbiomeMarker::normalize(tb_ec_cleaned, method = "CLR")
res.pca = PCA(tb_ec_clr, scale.unit = FALSE,graph = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(res.pca$eig[1:2,2], 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of MGS (CLR transfomation)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_EC_MGS_CLR.png", width = 8, height = 10, dpi = 300)
```
##PCA analysis (CLR transformation + imputation)
```{r}
tb_ec_clr1 = tb_ec_cleaned
tb_ec_clr1[tb_ec_clr1 == 0] = 0.01
tb_ec_clr1 = clr(tb_ec_clr1)
res.pca = PCA(tb_ec_clr1, scale.unit = FALSE,graph = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(res.pca$eig[1:2,2], 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of MGS (CLR transfomation + Imp)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_EC_MGS_CLR_imp.png", width = 8, height = 10, dpi = 300)
```
##PCoA analysis (Data raw)
```{r}
set.seed(38)
res.pcoa = ape::pcoa(vegan::vegdist(t(tb_ec_cleaned), method = "bray"))
permanova = vegan::adonis2(as.matrix(vegan::vegdist(t(tb_ec_cleaned)),method = "bray") ~ Category * time, data = meta, method = "bray", permutations = 999,sqrt.dist = FALSE, add = FALSE, by = "terms")
pv = permanova$`Pr(>F)`[1]
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of MGS (Data raw)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
    xlab(paste0("Axis.1 [", round(res.pcoa$values$Relative_eig[[1]] * 100, digits = 2), " %] ", "(pvalue=", pv, ")")) +
    ylab(paste0("Axis.2 [", round(res.pcoa$values$Relative_eig[[2]] * 100, digits = 2), " %]")) +
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCoA_EC_MGS.png", width = 8, height = 10, dpi = 300)
```
##PCoA analysis (CLR transformation)
```{r}
set.seed(38)
res.pcoa = ape::pcoa(vegan::vegdist(t(tb_ec_clr), method = "euclidean"))
permanova = vegan::adonis2(as.matrix(vegan::vegdist(t(tb_ec_clr)),method = "euclidean") ~ Category * time, data = meta, method = "euclidean", permutations = 999,sqrt.dist = FALSE, add = FALSE, by = "terms")
pv = permanova$`Pr(>F)`[1]
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of MGS (CLR transformation)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
    xlab(paste0("Axis.1 [", round(res.pcoa$values$Relative_eig[[1]] * 100, digits = 2), " %] ", "(pvalue=", pv, ")")) +
    ylab(paste0("Axis.2 [", round(res.pcoa$values$Relative_eig[[2]] * 100, digits = 2), " %]")) +
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCoA_EC_MGS_clr.png", width = 8, height = 10, dpi = 300)
```
##PCoA analysis (CLR transformation + imputation)
```{r}
set.seed(38)
res.pcoa = ape::pcoa(vegan::vegdist(t(tb_ec_clr1), method = "euclidean"))
permanova = vegan::adonis2(as.matrix(vegan::vegdist(t(tb_ec_clr1)),method = "euclidean") ~ Category * time, data = meta, method = "euclidean", permutations = 999,sqrt.dist = FALSE, add = FALSE, by = "terms")
pv = permanova$`Pr(>F)`[1]
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of MGS (CLR transformation + IMP)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
    xlab(paste0("Axis.1 [", round(res.pcoa$values$Relative_eig[[1]] * 100, digits = 2), " %] ", "(pvalue=", pv, ")")) +
    ylab(paste0("Axis.2 [", round(res.pcoa$values$Relative_eig[[2]] * 100, digits = 2), " %]")) +
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCoA_EC_MGS_clr_imp.png", width = 8, height = 10, dpi = 300)
```
##PCoA analysis (VST)
```{r}
set.seed(38)
res.pcoa = ape::pcoa(vegan::vegdist(t(assay(vst_MGS)), method = "euclidean"))
permanova = vegan::adonis2(as.matrix(vegan::vegdist(t(assay(vst_MGS))),method = "euclidean") ~ Category * time, data = meta, method = "euclidean", permutations = 999,sqrt.dist = FALSE, add = FALSE, by = "terms")
pv = permanova$`Pr(>F)`[1]
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=10) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 6, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of MGS (VST)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  xlab(paste0("Axis.1 [", round(res.pcoa$values$Relative_eig[[1]] * 100, digits = 2), " %] ", "(pvalue=", pv, ")")) +
  ylab(paste0("Axis.2 [", round(res.pcoa$values$Relative_eig[[2]] * 100, digits = 2), " %]")) +
  guides(color = FALSE, fill = FALSE) +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCoA_EC_MGS_VST.png", width = 8, height = 8, dpi = 300)
```
##Differentially abundance analysis for Cropland
```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        EC_number = character())
tb_ec_cleaned = as.data.frame(tb_ec_cleaned)
for ( dataset in c("crop_0", "crop_3", "crop_51", "crop_125")){
  counts_matrix <- as.matrix(tb_ec_cleaned[,get(dataset)],rownames = TRUE)
  counts_matrix <- round(counts_matrix,0)
  sample_data <- meta[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ wheat_straw_amendment, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$EC_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, EC_number_class, by = "EC_number")
write_rds(res4_df, "Output/DA_EC_MGS_Crops.rds")
```
##Differentially abundance analysis for Grassland
```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        EC_number = character())
for ( dataset in c("grass_0", "grass_3", "grass_51", "grass_125")){
  counts_matrix <- as.matrix(tb_ec_cleaned[,get(dataset)],rownames = TRUE)
  counts_matrix <- round(counts_matrix,0)
  sample_data <- meta[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ wheat_straw_amendment, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$EC_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, EC_number_class, by = "EC_number")
write_rds(res4_df, "Output/DA_EC_MGS_Grass.rds")
```
##Differentially abundance analysis for Control (Land-use effect)
```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        EC_number = character())
tb_ec_cleaned = as.data.frame(tb_ec_cleaned)
for ( dataset in c("control_0", "control_3", "control_51", "control_125")){
  counts_matrix <- as.matrix(tb_ec_cleaned[,get(dataset)],rownames = TRUE)
  counts_matrix <- round(counts_matrix,0)
  sample_data <- meta[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ land_use, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$EC_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, EC_number_class, by = "EC_number")
write_rds(res4_df, "Output/DA_EC_MGS_Control_(land_use).rds")
```
##Differentially abundance analysis for Amended (Land use effect)
```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        EC_number = character())
tb_ec_cleaned = as.data.frame(tb_ec_cleaned)
for ( dataset in c("amended_0", "amended_3", "amended_51", "amended_125")){
  counts_matrix <- as.matrix(tb_ec_cleaned[,get(dataset)],rownames = TRUE)
  counts_matrix <- round(counts_matrix,0)
  sample_data <- meta[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ land_use, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$EC_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, EC_number_class, by = "EC_number")
write_rds(res4_df, "Output/DA_EC_MGS_Amended_(land_use).rds")
```
#Heatmap of EC
##Data transformation
```{r}
DA_EC_MGS = read_rds("Output/DA_EC_MGS_Crops.rds")
DA_EC = DA_EC_MGS$EC_number
count_MGS = assay(vst_MGS)[DA_EC,]
zscore_EC = t(scale(t(count_MGS), center = TRUE, scale = TRUE))
```
##Distance between the variables (intergroups, Control vs Amended)
```{r}
dist_a_vs_c = vegan::vegdist(t(count_MGS), method = "euclidean")
tree_a_vs_c = hclust(dist_a_vs_c, method = "complete")
cutree_a_vs_c = cutree(tree_a_vs_c, k = 7 )
table(cutree_a_vs_c)
```
##Distance between the observation (EC)
```{r}
dist_ec = vegan::vegdist(count_MGS, method = "euclidean")
tree_ec = hclust(dist_ec, method = "complete")
#cut into arbitrary clusters
cutree_ec = cutree(tree_ec, k = 5 )
table(cutree_ec)
```
##Heatmap <!--Voir avec Domitille-->
```{r}
anno_col = as.data.frame(meta[,c("Category","time")])
colnames(anno_col) = c("Groups","Time")
rownames(anno_col) = rownames(meta)
anno_col$Groups = as.factor(anno_col$Groups)
anno_col$Time = as.factor(anno_col$Time)
anno_row = EC_number_class
anno_row = dplyr::select(anno_row, -EC_number)
time_colors <- c(
  "0" = "#1f77b4",   
  "3" = "#ff7f0e",   
  "51" = "#2ca02c",  
  "125" = "#d62728"  
)
ec_colors <- c("Hydrolases" = "#E41A1C",  # Red
  "Isomerases" = "#377EB8",  # Blue
  "Ligases" = "#4DAF4A",  # Green
  "Lyases" = "#984EA3",  # Purple
  "Oxidoreductases" = "#FF7F00",  # Orange
  "Transferases" = "#A65628")  # Yellow)
vect_colours = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd", 
                "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2")
list_colors = list(Groups = vect_colours, Time = time_colors, EC_class = ec_colors)
pheatmap(zscore_EC,
            cluster_rows = tree_ec, cluster_cols = tree_a_vs_c,
              show_rownames = FALSE, 
              cutree_rows = 5, #9, #5
              cutree_cols = 7,
              annotation_colors = list_colors,
              annotation_col = anno_col, 
              annotation_row = anno_row)
```


Part II : _Analysis of COG abundance in the output of MGS_
Tidy up the table abundance of COG
```{r}
tb_cog = as.data.frame(tb_cog)
  rownames(tb_cog) = tb_cog[,1]
  tb_cog = tb_cog[,-1]
overlapping_COG = read_rds("Output/overlapping_COG.rds")
tb_cog_cleaned = round(tb_cog[overlapping_COG,],0)
```

```{r}
dds_MGS = DESeqDataSetFromMatrix(countData = tb_cog_cleaned,
                                  colData = meta,
                                  design = ~ Category + time, tidy = FALSE)
vst_COG_MGS = vst(dds_MGS,blind = FALSE) # variance stabilizing transformation
```

PCA for COG
```{r}
#PCA plot
pcaData = plotPCA(vst_COG_MGS, intgroup = c("Category", "time"), returnData = TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of COG abundance in output of MGS") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_COG_MGS.png", width = 8, height = 10, dpi = 300)
```
PCoA for COG with test permanova
```{r}
set.seed(38)
res.pcoa.COG = ape::pcoa(vegan::vegdist(t(assay(vst_COG_MGS)), method = "euclidean"))
permanova = vegan::adonis2(as.matrix(vegan::vegdist(t(assay(vst_COG_MGS))),method = "euclidean") ~ Category * time, data = meta, method = "euclidean", permutations = 999,sqrt.dist = FALSE, add = FALSE, by = "terms")
pv = permanova$`Pr(>F)`[1]
pcoaData = as.data.frame(res.pcoa.COG$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=10) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 6, alpha = 0.6)+
 ggtitle("PCoaA of COG abundance in output of MGS (VST)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  xlab(paste0("Axis.1 [", round(res.pcoa.COG$values$Relative_eig[[1]] * 100, digits = 2), " %] ", "(pvalue=", pv, ")")) +
  ylab(paste0("Axis.2 [", round(res.pcoa.COG$values$Relative_eig[[2]] * 100, digits = 2), " %]")) +
  guides(color = FALSE, fill = FALSE) +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCoA_COG_MGS_VST.png", width = 8, height = 8, dpi = 300)
```
DAG analysis about COG for Cropland
```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        COG_number = character())
tb_cog_cleaned = as.data.frame(tb_cog_cleaned)
for ( dataset in c("crop_0", "crop_3", "crop_51", "crop_125")){
  counts_matrix <- as.matrix(tb_cog_cleaned[,get(dataset)],rownames = TRUE)
  sample_data <- meta[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ wheat_straw_amendment, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$COG_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, COG_anno, by = "COG_number")
write_rds(res4_df, "Output/DA_COG_MGS_Crops.rds")
```
DAG analysis about COG for Grassland
```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        COG_number = character())
tb_cog_cleaned = as.data.frame(tb_cog_cleaned)
for ( dataset in c("grass_0", "grass_3", "grass_51", "grass_125")){
  counts_matrix <- as.matrix(tb_cog_cleaned[,get(dataset)],rownames = TRUE)
  sample_data <- meta[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ wheat_straw_amendment, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$COG_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, COG_anno, by = "COG_number")
write_rds(res4_df, "Output/DA_COG_MGS_Grass.rds")
```
Selection des COG significatives pour heatmap + Génération de la matrix des comptages
```{r}
DA_COG_MGS_Crop = read_rds("Output/DA_COG_MGS_Crops.rds")
DA_COG_MGS_Grass = read_rds("Output/DA_COG_MGS_Grass.rds")
DA_COG_P2_Crop = read_rds("Output/DA_COG_P2_Crops.rds")
DA_COG_P2_Grass = read_rds("Output/DA_COG_P2_Grass.rds")
COG_Crop = unique(DA_COG_MGS_Crop$COG_number)
COG_Grass = unique(DA_COG_MGS_Grass$COG_number)
DA_COG = unique(c(COG_Crop, COG_Grass))
DA_COG_commun = intersect(unique(c(DA_COG_MGS_Crop$COG_number,DA_COG_MGS_Grass$COG_number)), 
                          unique(c(DA_COG_P2_Crop$COG_number,DA_COG_P2_Grass$COG_number)))
count_COG_MGS = assay(vst_COG_MGS)[DA_COG_commun,]
zscore_COG = t(scale(t(count_COG_MGS), center = TRUE, scale = TRUE))
```
##Distance between the variables (intergroups, Control vs Amended)
```{r}
#dist_a_vs_c = vegan::vegdist(t(count_COG_MGS), method = "eucleadean") #changé de eucleadean
dist_a_vs_c = as.dist(1-cor(count_COG_MGS,use = "everything", method = "pearson")) #changé de eucleadean
dist_a_vs_c = as.matrix(dist_a_vs_c)
dist_a_vs_c = dist_a_vs_c[rowSums(is.na(dist_a_vs_c)) == 0,colSums(is.na(dist_a_vs_c)) == 0,drop = FALSE]
dist_a_vs_c = as.dist(dist_a_vs_c)
tree_a_vs_c = hclust(dist_a_vs_c, method = "complete")
cutree_a_vs_c = cutree(tree_a_vs_c, k = 6 )
table(cutree_a_vs_c)
```
##Distance between the observation (COG)
```{r}
#dist_cog = vegan::vegdist(count_COG_MGS, method = "eucleadean") #changé de eucleadean
dist_cog = as.dist(1-cor(t(count_COG_MGS), use = "everything", method = "pearson")) #changé de eucleadean
dist_cog = as.matrix(dist_cog)
dist_cog = dist_cog[rowSums(is.na(dist_cog)) == 0,colSums(is.na(dist_cog)) == 0,drop = FALSE]
dist_cog = as.dist(dist_cog)
tree_cog = hclust(dist_cog, method = "complete")
#cut into arbitrary clusters
cutree_cog = cutree(tree_cog, k = 11 )
table(cutree_cog)
```
Heatmap for COG
```{r}
anno_col = as.data.frame(meta[,c("Category","time")])
colnames(anno_col) = c("Groups","Time")
rownames(anno_col) = rownames(meta)
anno_col$Groups = as.factor(anno_col$Groups)
anno_col$Time = as.factor(anno_col$Time)
anno_row = COG_anno[rownames(zscore_COG),]
anno_row$annotation = paste0(anno_row$COG_letter,":[",anno_row$COG_category,"]")
time_colors <- c(
  "0" = "#1f77b4",   
  "3" = "#ff7f0e",   
  "51" = "#353835",  
  "125" = "#d62728")
vect_colours = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd", 
                "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2")
hb = HeatmapAnnotation(which = "column",
                       Groups = anno_col$Groups,
                       col = list(Groups = vect_colours),
                       show_legend = TRUE,
                       annotation_name_side = "right",
                       annotation_name_gp = gpar(fontsize = 10, fontface = "bold"),
                       annotation_legend_param = list(direction = "vertical", title_gp = gpar(fontsize = 10, fontface = "bold")))
ha = rowAnnotation(COG = anno_row$annotation,
                   col = list(COG = c("J:[Translation, ribosomal structure and biogenesis]" = "#ff0000", 
"A:[RNA processing and modification]" = "#c2af58", "K:[Transcription]" = "#ff9900", 
"L:[Replication, recombination and repair]" = "#ffff00", "B:[Chromatin structure and dynamics]" = "#ffc600", 
"D:[Cell cycle control, cell division, chromosome partitioning]" = "#99ff00", 
"Y:[Nuclear structure]" = "#493126", "V:[Defense mechanisms]" = "#ff008a", 
"T:[Signal transduction mechanisms]" = "#0000ff", "M:[Cell wall/membrane/envelope biogenesis]" = "#9ec928", 
"N:[Cell motility]" = "#006633", "Z:[Cytoskeleton]" = "#660099", 
"W:[Extracellular structures]" = "#336699", "U:[Intracellular trafficking, secretion, and vesicular transport]" = "#33cc99", 
"O:[Posttranslational modification, protein turnover, chaperones]" = "#00ffff", 
"X:[Mobilome: prophages, transposons]" = "#9900ff", "C:[Energy production and conversion]" = "#805642", 
"G:[Carbohydrate transport and metabolism]" = "#ff00ff", "E:[Amino acid transport and metabolism]" = "#99334d", 
"F:[Nucleotide transport and metabolism]" = "#727dcc", "H:[Coenzyme transport and metabolism]" = "#5c5a1b", 
"I:[Lipid transport and metabolism]" = "#0099ff", "P:[Inorganic ion transport and metabolism]" = "#ffcc99", 
"Q:[Secondary metabolites biosynthesis, transport and catabolism]" = "#ff9999", 
"R:[General function prediction only]" = "#d6aadf", "S:[Function unknown]" = "#FFFFFF")),
                    Cluster = anno_block(gp = gpar(fill = c("1" = "#8DD3C7", "2" = "#FFFFB3", "3" = "#BEBADA", 
                                            "4" = "#FB8072", "5" = "#80B1D3", "6" = "#FDB462", 
                                            "7" = "#B3DE69", "8" = "#FCCDE5", "9" = "#D9D9D9",
                                            "10" = "#BC80BD", "11" = "#CCEBC5")),labels = c(1:11), labels_gp = gpar(fontsize = 8)))


p = pheatmap(zscore_COG, name = "Z-score",
            cluster_rows = tree_cog, cluster_cols = tree_a_vs_c,
              cutree_rows = 11, #9, #5
              cutree_cols = 6,
              top_annotation = hb,
              #right_annotation = ha ,
              show_rownames = FALSE)
pdf("Output/Heatmap_COG_MGS.pdf", width = 10, height = 16)
draw(p,align_heatmap_legend = "heatmap_top",merge_legend = TRUE, heatmap_legend_side = "right", annotation_legend_side = "right")
dev.off()
```
Regroupement des COGs en fonction de leur catégorie
```{r}
cat_count_COG = as.data.frame(count_COG_MGS)
cat_count_COG$COG_number = rownames(cat_count_COG)
cat_count_COG = left_join(cat_count_COG, COG_anno[,c(1,4)], by = "COG_number")
cat_count_COG = cat_count_COG %>% dplyr::select(-COG_number)
cat_count_COG$COG_letter = factor(cat_count_COG$COG_letter, levels = c("Z","Y","X","W","V","U","T","S","R","Q","P","O","N","M","L","K","J","I","H","G","F","E","D","C","B","A"))
cat_count_COG = aggregate(. ~ COG_letter, data = cat_count_COG, FUN = sum)
rownames(cat_count_COG) = cat_count_COG$COG_letter
cat_count_COG = cat_count_COG %>% dplyr::select(-COG_letter)
```

```{r}
bar_plot_cog = list()
for ( time in c(0, 3, 51, 125)){
  df_count = cat_count_COG[,c(get(paste0("crop_", time)), get(paste0("grass_", time)))]
  df_dif <- data.frame(
      dif_C_1 = df_count[, 4] - df_count[, 1],
      dif_C_2 = df_count[, 5] - df_count[, 2],
      dif_C_3 = df_count[, 6] - df_count[, 3],
      dif_G_1 = df_count[, 10] - df_count[, 7],
      dif_G_2 = df_count[, 11] - df_count[, 8],
      dif_G_3 = df_count[, 12] - df_count[, 9]
    )
  names(df_dif) = c("dif_C_1", "dif_C_2","dif_C_3","dif_G_1","dif_G_2","dif_G_3")
  rownames(df_dif) = rownames(df_count)
  df_dif$mean_C = rowMeans(df_dif[,1:3])
  df_dif$mean_G = rowMeans(df_dif[,4:6])
  df_dif$sd_C = apply(df_dif[,1:3], 1, sd)
  df_dif$sd_G = apply(df_dif[,4:6], 1, sd)
  df_dif$COG_letter = as.factor(rownames(df_dif))
  df_plot = df_dif[,7:11]
  df_plot = df_plot %>% pivot_longer(cols = 1:4,names_to = c(".value", "type"),names_sep = "_")
  df_plot$type = factor(df_plot$type, levels = c("G","C"))
  bar_plot_cog[[paste0("T",time)]] = ggplot(df_plot, aes(x = mean, y = COG_letter, fill = type)) +
    geom_bar(position = "dodge", stat = "identity") +
    geom_errorbar(aes(y = COG_letter, xmin = mean - sd, xmax = mean + sd), width = 0.2, position = position_dodge(0.9)) +
    xlim(-7, 7) +
    xlab(NULL) +
    ylab(NULL) +
    ggtitle(paste0("Difference between Amended and Control at "," (T",time,") WGS's outcome")) +
    scale_fill_manual(values = c("C" = "#fbbb3f", "G" = "#2abf2a"), labels = c("C" = "Cropland","G" = "Grassland")) +
    geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
    theme_bw() +
    theme(plot.title = element_text(size = 12, hjust =0.5, face = "bold"),
          axis.text.y = element_text(size = 12, face = "bold"),
          legend.position = "none",
          legend.title=element_blank(),
          legend.key.size = unit(1, 'cm'),
          legend.text = element_text(size=8))
}
ggsave(bar_plot_cog[[1]], file = "Output/Barplot_COG_MGS_T0.png", width = 6, height = 8, dpi = 300)
ggsave(bar_plot_cog[[2]], file = "Output/Barplot_COG_MGS_T3.png", width = 6, height = 8, dpi = 300)
ggsave(bar_plot_cog[[3]], file = "Output/Barplot_COG_MGS_T51.png", width = 6, height = 8, dpi = 300)
ggsave(bar_plot_cog[[4]], file = "Output/Barplot_COG_MGS_T125.png", width = 6, height = 8, dpi = 300)
```
```{r}
COG_DA_a_vs_c = anno_row
COG_DA_a_vs_c$cluster = cutree_cog
COG_DA_a_vs_c[COG_DA_a_vs_c$cluster == 1,"cluster"] = "0.1"
COG_DA_a_vs_c[COG_DA_a_vs_c$cluster == 2,"cluster"] = "0.2"
COG_DA_a_vs_c[COG_DA_a_vs_c$cluster == 3,"cluster"] = "0.3"
COG_DA_a_vs_c[COG_DA_a_vs_c$cluster == 4,"cluster"] = "0.4"
COG_DA_a_vs_c[COG_DA_a_vs_c$cluster == 5,"cluster"] = "0.5"
COG_DA_a_vs_c[COG_DA_a_vs_c$cluster == 6,"cluster"] = "1.0"
COG_DA_a_vs_c[COG_DA_a_vs_c$cluster == 7,"cluster"] = "0.8"
COG_DA_a_vs_c[COG_DA_a_vs_c$cluster == 8,"cluster"] = "1.1"
COG_DA_a_vs_c[COG_DA_a_vs_c$cluster == 9,"cluster"] = "0.9"
COG_DA_a_vs_c[COG_DA_a_vs_c$cluster == 10,"cluster"] = "0.6"
COG_DA_a_vs_c[COG_DA_a_vs_c$cluster == 11,"cluster"] = "0.7"

COG_DA_a_vs_c$cluster = as.numeric(COG_DA_a_vs_c$cluster)
COG_DA_a_vs_c$cluster = COG_DA_a_vs_c$cluster * 10
COG_DA_a_vs_c$cluster = as.factor(COG_DA_a_vs_c$cluster)
```

```{r}
df_count_COG_MGS = as.data.frame(tb_cog[DA_COG_commun,])
df_count_COG_MGS$cluster = COG_DA_a_vs_c$cluster
df_count_COG_MGS = arrange(df_count_COG_MGS, cluster)
COG_DA_a_vs_c = arrange(COG_DA_a_vs_c, cluster)
write_rds(df_count_COG_MGS, "Output/DA_count_COG_MGS.rds")
write_delim(COG_DA_a_vs_c, "Output/DA_COG_MGS.tsv", col_names = TRUE, delim = "\t")
```
Split df for each cluster
```{r}
list_cluster = list()
for(n in c(1:11)){
  list_cluster[[paste0("cluster_", n)]] = df_count_COG_MGS[df_count_COG_MGS$cluster == n,]
  }
```

```{r}
plot_da_cog = list()
COG_DA_a_vs_c$COG_letter = factor(COG_DA_a_vs_c$COG_letter, levels = c("A","B","C","D","E","F","G","H","I","J","K","L","M","N","O","P","Q","R","S","T","U","V","W","X","Y","Z"))
for (n in (1:11)){
    for ( time in c(0, 3, 51, 125)){
  df_cluster = list_cluster[[n]][,c(get(paste0("crop_", time)), get(paste0("grass_", time)))]
  df_plot_cluster <- data.frame(
      dif_C_1 = df_cluster[, 4] - df_cluster[, 1],
      dif_C_2 = df_cluster[, 5] - df_cluster[, 2],
      dif_C_3 = df_cluster[, 6] - df_cluster[, 3],
      dif_G_1 = df_cluster[, 10] - df_cluster[, 7],
      dif_G_2 = df_cluster[, 11] - df_cluster[, 8],
      dif_G_3 = df_cluster[, 12] - df_cluster[, 9]
    )
  names(df_plot_cluster) = c("dif_C_1", "dif_C_2","dif_C_3","dif_G_1","dif_G_2","dif_G_3")
  rownames(df_plot_cluster) = rownames(df_cluster)
  df_plot_cluster$anno = factor(COG_DA_a_vs_c[COG_DA_a_vs_c$cluster == n,"COG_letter"],levels = c("Z","Y","X","W","V","U","T","S","R","Q","P","O","N","M","L","K","J","I","H","G","F","E","D","C","B","A"))
  df_plot_cluster = df_plot_cluster %>% pivot_longer(cols = 1:6, names_to = "sample", values_to = "abundance")
  df_plot_cluster$sample = rep(c("Crop","Grass"), each = 3, times = nrow(df_plot_cluster)/6)
  df_plot_cluster$sample = factor(df_plot_cluster$sample, levels = c("Grass","Crop"))
  plot_da_cog[[paste0("cluster_",n,"_T",time)]] = ggplot(df_plot_cluster, aes(x = abundance, y = anno, fill = sample, color = anno)) +
    geom_bar(position = "dodge", stat = "summary", fun = "mean" ) +
    xlab(NULL) +
    ylab(NULL) +
    ggtitle(paste0("Cluster ", n, " (T",time,")")) +
    scale_fill_manual(values = c("Crop" = "#fbbb3f", "Grass" = "#2abf2a")) +
    scale_color_manual(values = c("J" = "#ff0000", 
"A" = "#c2af58", "K" = "#ff9900", 
"L" = "#ffff00", "B" = "#ffc600", 
"D" = "#99ff00", 
"Y" = "#493126", "V" = "#ff008a", 
"T" = "#0000ff", "M" = "#9ec928", 
"N" = "#006633", "Z" = "#660099", 
"W" = "#336699", "U" = "#33cc99", 
"O" = "#00ffff", 
"X" = "#9900ff", "C" = "#805642", 
"G" = "#ff00ff", "E" = "#99334d", 
"F" = "#727dcc", "H" = "#5c5a1b", 
"I" = "#0099ff", "P" = "#ffcc99", 
"Q" = "#ff9999", 
"R" = "#d6aadf", "S" = "#FFFFFF")) +
    geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
    theme_bw() +
    guides(color = "none") +
    theme(plot.title = element_text(size = 8, hjust =0.5, face = "bold"),
          axis.text.y = element_text(size = 8, face = "bold"),
          legend.title=element_blank(),
          legend.key.size = unit(0.5, 'cm'),
          legend.text = element_text(size=3))
  }
}
```

```{r}
plot_grid = list()
  for (i in c(1,5,9,13,17,21,25,29,33,37,41)){
    plot_grid[[paste0("Cluster ",i)]] = ggarrange(plot_da_cog[[i]],plot_da_cog[[i+1]],plot_da_cog[[i+2]],plot_da_cog[[i+3]], ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
  }
```

```{r}
tb_pathways = as.data.frame(tb_pathways)
rownames(tb_pathways) = tb_pathways[,1]
tb_pathways = tb_pathways[,-1]
```

```{r}
pathways_dds = round(tb_pathways,0)
dds_MGS = DESeqDataSetFromMatrix(countData = pathways_dds,
                                  colData = meta,
                                  design = ~ Category + time, tidy = FALSE)
vst_pathways_MGS = varianceStabilizingTransformation(dds_MGS,blind = FALSE) # variance stabilizing transformation
```

```{r}
set.seed(38)
res.pcoa = ape::pcoa(vegan::vegdist(t(assay(vst_pathways_MGS)), method = "euclidean"))
permanova = vegan::adonis2(as.matrix(vegan::vegdist(t(assay(vst_pathways_MGS))),method = "euclidean") ~ Category * time, data = meta, method = "euclidean", permutations = 999,sqrt.dist = FALSE, add = FALSE, by = "terms")
pv = permanova$`Pr(>F)`[1]
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of metabolics pathways abundance in output of MGS (VST)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  xlab(paste0("Axis.1 [", round(res.pcoa$values$Relative_eig[[1]] * 100, digits = 2), " %] ", "(pvalue=", pv, ")")) +
  ylab(paste0("Axis.2 [", round(res.pcoa$values$Relative_eig[[2]] * 100, digits = 2), " %]")) +
  guides(color = guide_legend(title = "Groups")) +
  theme_bw()+ 
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCoA_metacyc_MGS_VST.png", width = 8, height = 10, dpi = 300)
```

```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        pathways = character())
tb_pathways = as.data.frame(tb_pathways)
for ( dataset in c("crop_0","crop_3", "crop_51", "crop_125")){              
  counts_matrix <- as.matrix(tb_pathways[,get(dataset)],rownames = TRUE)
  counts_matrix <- round(counts_matrix,0)
  sample_data <- meta[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ wheat_straw_amendment, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$pathways = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
write_rds(res4_df, "Output/DA_pathways_MGS_Crops.rds")
```

```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        pathways = character())
tb_pathways = as.data.frame(tb_pathways)
for ( dataset in c("grass_0","grass_3", "grass_51", "grass_125")){              
  counts_matrix <- as.matrix(tb_pathways[,get(dataset)],rownames = TRUE)
  counts_matrix <- round(counts_matrix,0)
  sample_data <- meta[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ wheat_straw_amendment, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$pathways = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
write_rds(res4_df, "Output/DA_pathways_MGS_Grass.rds")
```

```{r}
DA_pathways_MGS_Crop = read_rds("Output/DA_pathways_MGS_Crops.rds")
DA_pathways_MGS_Grass = read_rds("Output/DA_pathways_MGS_Grass.rds")
DA_pathways_signif = read_rds("DA_metacyc_signif.rds")
pathways_Crop = unique(DA_pathways_MGS_Crop$pathways)
pathways_Grass = unique(DA_pathways_MGS_Grass$pathways)
DA_pathways = unique(c(pathways_Crop, pathways_Grass))
count_pathways_MGS = assay(vst_pathways_MGS)[DA_pathways_signif,]
zscore_pathways = t(scale(t(count_pathways_MGS), center = TRUE, scale = TRUE))
```

##Distance between the variables (intergroups, Control vs Amended)
```{r}
#dist_a_vs_c = vegan::vegdist(t(count_pathways_MGS), method = "euclidean")
dist_a_vs_c = as.dist(1-cor(count_pathways_MGS,use = "everything", method = "pearson")) #changé de eucleadean
dist_a_vs_c = as.matrix(dist_a_vs_c)
dist_a_vs_c = dist_a_vs_c[rowSums(is.na(dist_a_vs_c)) == 0,colSums(is.na(dist_a_vs_c)) == 0,drop = FALSE]
dist_a_vs_c = as.dist(dist_a_vs_c)
tree_a_vs_c = hclust(dist_a_vs_c, method = "complete")
cutree_a_vs_c = cutree(tree_a_vs_c, k = 4 )
table(cutree_a_vs_c)
```
##Distance between the observation (pathways)
```{r}
#dist_metacyc = vegan::vegdist(count_pathways_MGS, method = "euclidean")
dist_metacyc = as.dist(1-cor(t(count_pathways_MGS), use = "everything", method = "pearson")) #changé de eucleadean
dist_metacyc = as.matrix(dist_metacyc)
dist_metacyc = dist_metacyc[rowSums(is.na(dist_metacyc)) == 0,colSums(is.na(dist_metacyc)) == 0,drop = FALSE]
dist_metacyc = as.dist(dist_metacyc)
tree_metacyc = hclust(dist_metacyc, method = "complete")
#cut into arbitrary clusters
cutree_metacyc = cutree(tree_metacyc, k = 5)
table(cutree_metacyc)
```

```{r}
anno_col = as.data.frame(meta[,c("Category","time")])
colnames(anno_col) = c("Groups","Time")
rownames(anno_col) = rownames(meta)
anno_col$Groups = as.factor(anno_col$Groups)
anno_col$Time = as.factor(anno_col$Time)
time_colors <- c(
  "0" = "#1f77b4",   
  "3" = "#ff7f0e",   
  "51" = "#2ca02c",  
  "125" = "#d62728")
vect_colours = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd", 
                "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2")
hb = HeatmapAnnotation(which = "column",
                       Groups = anno_col$Groups,
                       #Time = anno_col$Time,
                       col = list(Groups = vect_colours, Time = time_colors),
                       show_legend = TRUE,
                       annotation_name_side = "right",
                       annotation_name_gp = gpar(fontsize = 10, fontface = "bold"),
                       annotation_legend_param = list(direction = "vertical", title_gp = gpar(fontsize = 10, fontface = "bold")))
p = pheatmap(zscore_pathways, name = "Z-score",
            cluster_rows = tree_metacyc, cluster_cols = tree_a_vs_c,
              cutree_rows = 5, #18, #6
              cutree_cols = 4,
              top_annotation = hb,
              show_rownames = TRUE)
```

```{r}
bar_plot = list()
for ( time in c(0, 3, 51, 125)){
  df_count = count_pathways_MGS[,c(get(paste0("crop_", time)), get(paste0("grass_", time)))]
  df_dif <- data.frame(
      dif_C_1 = df_count[, 4] - df_count[, 1],
      dif_C_2 = df_count[, 5] - df_count[, 2],
      dif_C_3 = df_count[, 6] - df_count[, 3],
      dif_G_1 = df_count[, 10] - df_count[, 7],
      dif_G_2 = df_count[, 11] - df_count[, 8],
      dif_G_3 = df_count[, 12] - df_count[, 9]
    )
  names(df_dif) = c("dif_C_1", "dif_C_2","dif_C_3","dif_G_1","dif_G_2","dif_G_3")
  rownames(df_dif) = rownames(df_count)
  df_dif$mean_C = rowMeans(df_dif[,1:3])
  df_dif$mean_G = rowMeans(df_dif[,4:6])
  df_dif$sd_C = apply(df_dif[,1:3], 1, sd)
  df_dif$sd_G = apply(df_dif[,4:6], 1, sd)
  df_dif$anno = as.factor(rownames(df_dif))
  df_plot = df_dif[,7:11]
  df_plot = df_plot %>% pivot_longer(cols = 1:4,names_to = c(".value", "type"),names_sep = "_")
  df_plot$type = factor(df_plot$type, levels = c("G","C"))
  bar_plot[[paste0("T",time)]] = ggplot(df_plot, aes(x = mean, y = anno, fill = type)) +
    geom_bar(position = "dodge", stat = "identity") +
    geom_errorbar(aes(y = anno, xmin = mean - sd, xmax = mean + sd), width = 0.2, position = position_dodge(0.9)) +
    xlim(-0.25, 0.25) +
    xlab(NULL) +
    ylab(NULL) +
    ggtitle(paste0("Difference between Amended and Control at "," (T",time,") WGS's outcome")) +
    scale_fill_manual(values = c("C" = "#fbbb3f", "G" = "#2abf2a"), labels = c("C" = "Cropland","G" = "Grassland")) +
    geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
    theme_bw() +
    theme(plot.title = element_text(size = 8, hjust =0.5, face = "bold"),
          axis.text.y = element_text(size = 8, face = "bold"),
          legend.title=element_blank(),
          legend.key.size = unit(1, 'cm'),
          legend.text = element_text(size=8))
}
```

```{r}
bar_plot_cog_i = list()
count_COG_MGS_i = count_COG_MGS[c("COG0814","COG2431","COG1114","COG4533","COG3160","COG3711","COG1440","COG1447","COG4211"),]
for ( time in c(0, 3, 51, 125)){
  df_count = count_COG_MGS_i[,c(get(paste0("crop_", time)), get(paste0("grass_", time)))]
  df_dif <- data.frame(
      dif_C_1 = df_count[, 4] - df_count[, 1],
      dif_C_2 = df_count[, 5] - df_count[, 2],
      dif_C_3 = df_count[, 6] - df_count[, 3],
      dif_G_1 = df_count[, 10] - df_count[, 7],
      dif_G_2 = df_count[, 11] - df_count[, 8],
      dif_G_3 = df_count[, 12] - df_count[, 9]
    )
  names(df_dif) = c("dif_C_1", "dif_C_2","dif_C_3","dif_G_1","dif_G_2","dif_G_3")
  rownames(df_dif) = rownames(df_count)
  df_dif$mean_C = rowMeans(df_dif[,1:3])
  df_dif$mean_G = rowMeans(df_dif[,4:6])
  df_dif$sd_C = apply(df_dif[,1:3], 1, sd)
  df_dif$sd_G = apply(df_dif[,4:6], 1, sd)
  df_dif$COG_number = factor(rownames(df_dif),levels = c("COG0814","COG2431","COG1114","COG4533","COG3160","COG3711","COG1440","COG1447","COG4211"))
  df_plot = df_dif[,7:11]
  df_plot = df_plot %>% pivot_longer(cols = 1:4,names_to = c(".value", "type"),names_sep = "_")
  df_plot$type = factor(df_plot$type, levels = c("G","C"))
  bar_plot_cog_i[[paste0("T",time)]] = ggplot(df_plot, aes(x = mean, y = COG_number, fill = type)) +
    geom_bar(position = "dodge", stat = "identity") +
    geom_errorbar(aes(y = COG_number, xmin = mean - sd, xmax = mean + sd), width = 0.2, position = position_dodge(0.9)) +
    xlab(NULL) +
    ylab(NULL) +
    ggtitle(paste0("Difference between Amended and Control at "," (T",time,") WGS's outcome")) +
    scale_fill_manual(values = c("C" = "#fbbb3f", "G" = "#2abf2a"), labels = c("C" = "Cropland","G" = "Grassland")) +
    geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
    theme_bw() +
    theme(plot.title = element_text(size = 8, hjust =0.5, face = "bold"),
          axis.text.y = element_text(size = 8, face = "bold"),
          legend.title=element_blank(),
          legend.key.size = unit(1, 'cm'),
          legend.text = element_text(size=8))
}
```

```{r}
b = dudi.pca(t(assay(vst_COG_MGS)), scannf = FALSE, nf = 2)
a = dudi.pca(t(assay(vst_MGS)), scannf = FALSE, nf = 2)
coi = coinertia(b, a, scannf = FALSE, nf = 2)
randtest(coi, fixed=1)
```

