---
title: "FONCTIOMIC_MGS"
auteur: "Dinh-Phuc Hoang"
date: "2025-03-21"
---


```{r}
library(tidyverse)
library(dplyr)
library(phyloseq)
library(DESeq2)
library(pheatmap)
library(BiocParallel)
library(FactoMineR)
library(microbiomeMarker)
register(SnowParam(4))
```
Data loading
```{r}
tb_ec = read_tsv("Data FONCTIOMIC/counts_by_EC_number.norm.tsv")
meta = read_tsv("Data FONCTIOMIC/complete_metadata_table.tsv")
meta = as.data.frame(meta)
row.names(meta) = meta$sample_id
EC_number_class = read.table("description_mapfiles/EC_number_class.txt", header = TRUE, sep = "\t")
rownames(EC_number_class) = EC_number_class$EC_number
tb_cog = read_tsv("Data FONCTIOMIC/counts_by_COG.raw.tsv")
COG_anno = read_delim("description_mapfiles/cog_info.tsv.gz", delim = "\t", col_names = FALSE)
names(COG_anno) = c("COG_number","description")
```
**Data description**
* tb_ec: abundance of EC number in the output of MGS
* meta: metadata of samples
* EC_number_class: description of EC number
* tb_ec_cleaned: abundance of EC number in the output of MGS after cleaning
Part I : _Analysis of EC abundance in the output of MGS_
Tidy up the table abundance of EC
```{r}
tb_ec = as.data.frame(tb_ec)
row.names(tb_ec) = paste0("EC:", tb_ec$EC_number)
tb_ec_cleaned = tb_ec %>% filter(!str_detect(rownames(tb_ec), "-$"))
tb_ec_cleaned = tb_ec_cleaned[,-c(1,50)]
tb_ec_cleaned = as.matrix(tb_ec_cleaned)
```
PCA analysis (VST)
```{r}
meta$Category = as.factor(meta$Category)
meta$Category = relevel(meta$Category, ref = "Control Cropland")
meta$time = as.factor(meta$time)
meta$time = relevel(meta$time, ref = "0")
meta$wheat_straw_amendment = as.factor(meta$wheat_straw_amendment)
meta$wheat_straw_amendment = relevel(meta$wheat_straw_amendment, ref = "Control")
tb_ec_cleaned = round(tb_ec_cleaned,0)
dds_MGS = DESeqDataSetFromMatrix(countData = tb_ec_cleaned,
                                  colData = meta,
                                  design = ~ Category + time, tidy = FALSE)
vst_MGS = vst(dds_MGS,blind = FALSE) # variance stabilizing transformation
#PCA plot
pcaData = plotPCA(vst_MGS, intgroup = c("Category", "time"), returnData = TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of MGS") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_EC_MGS.png", width = 8, height = 10, dpi = 300)
```
PCA analysis (Raw data)
```{r}
res.pca = PCA(tb_ec_cleaned, scale.unit = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(100 * res.pca$eig[1:2] / sum(res.pca$eig), 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of MGS (Data raw)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_EC_MGS_raw.png", width = 8, height = 10, dpi = 300)
```
PCA analysis (CLR transformation of package microbiomeMarker)
```{r}
tb_ec_clr = microbiomeMarker::normalize(tb_ec_cleaned, method = "CLR")
res.pca = PCA(tb_ec_clr, scale.unit = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(100 * res.pca$eig[1:2] / sum(res.pca$eig), 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of MGS (CLR transfomation)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_EC_MGS_CLR.png", width = 8, height = 10, dpi = 300)
```
PCA analysis (CLR transformation + imputation)
```{r}
tb_ec_clr1 = tb_ec_cleaned
tb_ec_clr1[tb_ec_clr1 == 0] = 0.01
tb_ec_clr1 = clr(tb_ec_clr1)
res.pca = PCA(tb_ec_clr1, scale.unit = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(100 * res.pca$eig[1:2] / sum(res.pca$eig), 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of MGS (CLR transfomation + Imp)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_EC_MGS_CLR_imp.png", width = 8, height = 10, dpi = 300)
```
PCoA analysis (Data raw)
```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(tb_ec_cleaned), method = "bray"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of MGS (Data raw)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCoA_EC_MGS.png", width = 8, height = 10, dpi = 300)
```
PCoA analysis (CLR transformation)
```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(tb_ec_clr), method = "euclidean"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of MGS (CLR transformation)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCoA_EC_MGS_clr.png", width = 8, height = 10, dpi = 300)
```
PCoA analysis (CLR transformation + imputation)
```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(tb_ec_clr1), method = "euclidean"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of MGS (CLR transformation + IMP)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCoA_EC_MGS_clr_imp.png", width = 8, height = 10, dpi = 300)
```
PCoA analysis (VST)
```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(assay(vst_MGS)), method = "euclidean"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of MGS (VST)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCoA_EC_MGS_VST.png", width = 8, height = 10, dpi = 300)
```
Define of variables for subset matrix of EC
```{r}
for (groups in c("Control Cropland","Amended Cropland","Control Grassland","Amended Grassland")){
  for (time in c("0","3","51","125")){
  assign(paste0(gsub(" ","_",groups),"_",time), str_sort(rownames(meta[meta$Category == groups & meta$time == time,])))
  }
}
crop_0 = c(Control_Cropland_0,Amended_Cropland_0)
crop_3 = c(Control_Cropland_3,Amended_Cropland_3)
crop_51 = c(Control_Cropland_51,Amended_Cropland_51)
crop_125 = c(Control_Cropland_125,Amended_Cropland_125)
grass_0 = c(Control_Grassland_0,Amended_Grassland_0)
grass_3 = c(Control_Grassland_3,Amended_Grassland_3)
grass_51 = c(Control_Grassland_51,Amended_Grassland_51)
grass_125 = c(Control_Grassland_125,Amended_Grassland_125)
```
DAG analysis for Cropland
```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        EC_number = character())
tb_ec_cleaned = as.data.frame(tb_ec_cleaned)
for ( dataset in c("crop_0", "crop_3", "crop_51", "crop_125")){
  counts_matrix <- as.matrix(tb_ec_cleaned[,get(dataset)],rownames = TRUE)
  sample_data <- meta[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ wheat_straw_amendment, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$EC_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, EC_number_class, by = "EC_number")
write_rds(res4_df, "Output/DA_EC_MGS_Crops.rds")
```
DAG analysis for Grassland
```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        EC_number = character())
for ( dataset in c("grass_0", "grass_3", "grass_51", "grass_125")){
  counts_matrix <- as.matrix(tb_ec_cleaned[,get(dataset)],rownames = TRUE)
  sample_data <- meta[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ wheat_straw_amendment, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$EC_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, EC_number_class, by = "EC_number")
write_rds(res4_df, "Output/DA_EC_MGS_Grass.rds")
```
Heatmap of EC
Data transformation
```{r}
DA_EC_MGS = read_rds("Output/DA_EC_MGS_Crops.rds")
count_MGS = tb_ec_cleaned
count_MGS = microbiomeMarker::normalize(count_MGS, method = "CLR")
count_MGS = count_MGS %>% as.data.frame %>% filter(rownames(count_MGS) %in% DA_EC_MGS$EC_number)
```
Distance between the variables (intergroups, Control vs Amended)
```{r}
dist_a_vs_c = vegan::vegdist(t(count_MGS), method = "euclidean")
tree_a_vs_c = hclust(dist_a_vs_c, method = "complete")
cutree_a_vs_c = cutree(tree_a_vs_c, k = 4 )
table(cutree_a_vs_c)
```
Distance between the observation (EC)
```{r}
dist_ec = vegan::vegdist(count_MGS, method = "euclidean")
tree_ec = hclust(dist_ec, method = "complete")
#cut into arbitrary clusters
cutree_ec = cutree(tree_ec, k = 5 )
table(cutree_ec)
```
Heatmap ##voir avec Domitille
```{r}
anno_col = as.data.frame(meta[,c("Category","time")])
colnames(anno_col) = c("Groups","Time")
rownames(anno_col) = rownames(meta)
anno_col$Groups = as.factor(anno_col$Groups)
anno_col$Time = as.factor(anno_col$Time)
anno_row = EC_number_class
anno_row = dplyr::select(anno_row, -EC_number)
time_colors <- c(
  "0" = "#1f77b4",   
  "3" = "#ff7f0e",   
  "51" = "#2ca02c",  
  "125" = "#d62728"  
)
ec_colors <- c("Hydrolases" = "#E41A1C",  # Red
  "Isomerases" = "#377EB8",  # Blue
  "Ligases" = "#4DAF4A",  # Green
  "Lyases" = "#984EA3",  # Purple
  "Oxidoreductases" = "#FF7F00",  # Orange
  "Transferases" = "#A65628")  # Yellow)
vect_colours = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd", 
                "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2")
list_colors = list(Groups = vect_colours, Time = time_colors, EC_class = ec_colors)
pheatmap(count_MGS,
            cluster_rows = tree_ec, cluster_cols = tree_a_vs_c,
              show_rownames = FALSE, 
              cutree_rows = 5, #9, #5
              cutree_cols = 4,
              annotation_colors = list_colors,
              annotation_col = anno_col, 
              annotation_row = anno_row)
```
Part II : _Analysis of COG abundance in the output of MGS_
Tidy up the table abundance of COG
```{r}
tb_cog = as.data.frame(tb_cog)
  rownames(tb_cog) = tb_cog[,1]
  tb_cog = tb_cog[,-1]
```
PCA for COG
```{r}
tb_cog_cleaned = round(tb_cog,0)
dds_MGS = DESeqDataSetFromMatrix(countData = tb_cog_cleaned,
                                  colData = meta,
                                  design = ~ Category + time, tidy = FALSE)
vst_MGS = vst(dds_MGS,blind = FALSE) # variance stabilizing transformation
#PCA plot
pcaData = plotPCA(vst_MGS, intgroup = c("Category", "time"), returnData = TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of COG abundance in output of MGS") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_COG_MGS.png", width = 8, height = 10, dpi = 300)
```
DAG analysis about COG for Cropland
```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        COG_number = character())
tb_cog_cleaned = as.data.frame(tb_cog_cleaned)
for ( dataset in c("crop_0", "crop_3", "crop_51", "crop_125")){
  counts_matrix <- as.matrix(tb_cog_cleaned[,get(dataset)],rownames = TRUE)
  sample_data <- meta[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ wheat_straw_amendment, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$COG_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, COG_anno, by = "COG_number")
write_rds(res4_df, "Output/DA_COG_MGS_Crops.rds")
```
DAG analysis about COG for Grassland
```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        COG_number = character())
tb_cog_cleaned = as.data.frame(tb_cog_cleaned)
for ( dataset in c("grass_0", "grass_3", "grass_51", "grass_125")){
  counts_matrix <- as.matrix(tb_cog_cleaned[,get(dataset)],rownames = TRUE)
  sample_data <- meta[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ wheat_straw_amendment, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$COG_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, COG_anno, by = "COG_number")
write_rds(res4_df, "Output/DA_COG_MGS_Grass.rds")
```