---
title: "FONCTIOMIC_P2"
auteur: "Dinh-Phuc Hoang"
date: "2025-03-21"
---


```{r}
library(tidyverse)
library(dplyr)
library(phyloseq)
library(DESeq2)
library(pheatmap)
library(BiocParallel)
library(FactoMineR)
library(microbiomeMarker)
register(SnowParam(4))
```
Data loading
```{r}
meta = read_delim("Data FONCTIOMIC/03_dimimos_terrain.csv",delim = ",",col_names = TRUE)
p2_EC = read_rds("Data FONCTIOMIC/picrust2out_ver2/p2_EC.rds")
p2_EC = p2_EC[,str_sort(colnames(p2_EC),decreasing = F)]
EC_number_class = read.table("description_mapfiles/EC_number_class.txt", header = TRUE, sep = "\t")
p2_COG = read_rds("Data FONCTIOMIC/picrust2out_ver2/p2_COG.rds")
p2_COG = p2_COG[,str_sort(colnames(p2_COG),decreasing = F)]
rownames(EC_number_class) = EC_number_class$EC_number
COG_anno = read_delim("description_mapfiles/cog_info.tsv.gz", delim = "\t", col_names = FALSE)
names(COG_anno) = c("COG_number","description")
```
**Data description**
* p2_EC: abundance of EC number in the output of PICRUSt2
* meta: metadata of samples
* EC_number_class: description of EC number
* p2_EC_cleaned: abundance of EC number in the output of PICRUSt2 after cleaning
* meta_cleaned: cleaned metadata of samples
Part I : _Analysis of EC abundance in the output of P2_
Tidy up the metadata of samples
```{r}
meta = as.data.frame(meta)
rownames(meta) = meta$sample
meta$sample = as.factor(meta$sample)
meta$sample = factor(meta$sample, levels = c(sort(levels(meta$sample), decreasing = F)))
meta_cleaned = meta[colnames(p2_EC),]
meta_cleaned$groupe2 = paste(meta_cleaned$bloc,meta_cleaned$modality,sep = "_")
meta_cleaned = meta_cleaned %>% filter(campagne %in% c("0","1","4","6"))
meta_cleaned$campagne = as.factor(meta_cleaned$campagne)
meta_cleaned$campagne = relevel(meta_cleaned$campagne, ref = "0")
meta_cleaned = meta_cleaned %>% mutate(time = fct_recode(campagne, 
"0" = "0", "3" = "1", "51" = "4", "125" = "6")) %>% mutate(groups = fct_recode(groupe2, 
"Control Cropland" = "crops_control", "Amended Cropland" = "crops_residues", 
"Control Grassland" = "grasslands_control", "Amended Grassland" = "grasslands_residues"))
```
Tidy up the table abundance of EC
```{r}
p2_EC_cleaned = p2_EC[,rownames(meta_cleaned)]
```
PCA analysis (VST)
```{r}
meta_cleaned$groups = as.factor(meta_cleaned$groups)
meta_cleaned$groups = relevel(meta_cleaned$groups, ref = "Control Cropland")
meta_cleaned$time = as.factor(meta_cleaned$time)
meta_cleaned$time = relevel(meta_cleaned$time, ref = "0")
p2_EC_cleaned = round(p2_EC_cleaned,0)
dds_P2 = DESeqDataSetFromMatrix(countData = p2_EC_cleaned,
                                  colData = meta_cleaned,
                                  design = ~ groups + time, tidy = FALSE)
vst_P2 = vst(dds_P2,blind = FALSE) # variance stabilizing transformation
pcaData = plotPCA(vst_P2, intgroup = c("groups", "time"), returnData = TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(x = PC1, y = PC2, color = groups)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of PICRUSt2") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_EC_P2.png", width = 8, height = 10, dpi = 300)
```
PCA analysis (Raw data)
```{r}
res.pca = PCA(p2_EC_cleaned, scale.unit = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta_cleaned[,c("groups","time","sample")], join_by(sample == sample), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(100 * res.pca$eig[1:2] / sum(res.pca$eig), 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = groups)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of P2 (Data raw)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_EC_P2_raw.png", width = 8, height = 10, dpi = 300)
```
PCA analysis (CLR transformation of package microbiomeMarker)
```{r}
p2_EC_clr = microbiomeMarker::normalize(p2_EC_cleaned, method = "CLR")
res.pca = PCA(p2_EC_clr, scale.unit = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta_cleaned[,c("groups","time","sample")], join_by(sample == sample), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(100 * res.pca$eig[1:2] / sum(res.pca$eig), 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = groups)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of P2 (CLR transfomation)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_EC_P2_CLR.png", width = 8, height = 10, dpi = 300)
```
PCA analysis (CLR transformation + imputation)
```{r}
p2_EC_clr1 = p2_EC_cleaned
p2_EC_clr1[p2_EC_clr1 == 0] = 0.01
p2_EC_clr1 = clr(p2_EC_clr1)
res.pca = PCA(p2_EC_clr1, scale.unit = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta_cleaned[,c("groups","time","sample")], join_by(sample == sample), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(100 * res.pca$eig[1:2] / sum(res.pca$eig), 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = groups)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of P2 (CLR transfomation + Imp)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_EC_P2_CLR_imp.png", width = 8, height = 10, dpi = 300)
```
PCoA analysis (Data raw)
```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(p2_EC_cleaned), method = "bray"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta_cleaned[,c("groups","time","sample")], join_by(sample == sample), )
pcoaData.centroids = pcoaData %>% group_by(groups, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("groups", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=groups), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = groups))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=groups), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of P2 (Data raw)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCoA_EC_P2.png", width = 8, height = 10, dpi = 300)
```
PCoA analysis (CLR transformation)
```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(p2_EC_clr), method = "euclidean"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta_cleaned[,c("groups","time","sample")], join_by(sample == sample), )
pcoaData.centroids = pcoaData %>% group_by(groups, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("groups", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=groups), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = groups))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=groups), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of P2 (CLR transformation)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCoA_EC_P2_CLR.png", width = 8, height = 10, dpi = 300)
```
PCoA analysis (CLR transformation + imputation)
```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(p2_EC_clr1), method = "euclidean"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta_cleaned[,c("groups","time","sample")], join_by(sample == sample), )
pcoaData.centroids = pcoaData %>% group_by(groups, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("groups", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=groups), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = groups))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=groups), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of P2 (CLR transformation + imp)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCoA_EC_P2_CLR_imp.png", width = 8, height = 10, dpi = 300)
```
PCoA analysis (VST)
```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(assay(vst_P2)), method = "euclidean"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta_cleaned[,c("groups","time","sample")], join_by(sample == sample), )
pcoaData.centroids = pcoaData %>% group_by(groups, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("groups", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=groups), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = groups))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=groups), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of P2 (VST)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCoA_EC_P2_VST.png", width = 8, height = 10, dpi = 300)
```
Define of variables for subset matrix of EC
```{r}
for (groups in c("Control Cropland","Amended Cropland","Control Grassland","Amended Grassland")){
  for (time in c("0","3","51","125")){
  assign(paste0(gsub(" ","_",groups),"_",time), str_sort(rownames(meta_cleaned[meta_cleaned$groups == groups & meta_cleaned$time == time,])))
  }
}
P2_crop_0 = c(Control_Cropland_0,Amended_Cropland_0)
P2_crop_3 = c(Control_Cropland_3,Amended_Cropland_3) # Erratum : manque de la variable Amended_Cropland_3 : C6-R-T1
P2_crop_51 = c(Control_Cropland_51,Amended_Cropland_51)
P2_crop_125 = c(Control_Cropland_125,Amended_Cropland_125)
P2_grass_0 = c(Control_Grassland_0,Amended_Grassland_0)
P2_grass_3 = c(Control_Grassland_3,Amended_Grassland_3)
P2_grass_51 = c(Control_Grassland_51,Amended_Grassland_51)
P2_grass_125 = c(Control_Grassland_125,Amended_Grassland_125)# Erratum : manque de la variable Control_Grassland_125 : P1-C-T6 et P5-C-T6
```
DAG analysis for Cropland
```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        EC_number = character())
p2_EC_cleaned = as.data.frame(p2_EC_cleaned)
for ( dataset in c("P2_crop_0","P2_crop_3", "P2_crop_51", "P2_crop_125")){              # Erratum : manque de la variable P2_crop_3
  counts_matrix <- as.matrix(p2_EC_cleaned[,get(dataset)],rownames = TRUE)
  sample_data <- meta_cleaned[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ modality, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$EC_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, EC_number_class, by = "EC_number")
write_rds(res4_df, "Output/DA_EC_P2_Crops.rds")
```
DAG analysis for Grassland
```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        EC_number = character())
for ( dataset in c("P2_grass_0", "P2_grass_3", "P2_grass_51","P2_grass_125")){             # Erratum : manque de la variable P2_grass_125
  counts_matrix <- as.matrix(p2_EC_cleaned[,get(dataset)],rownames = TRUE)
  sample_data <- meta_cleaned[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ modality, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$EC_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, EC_number_class, by = "EC_number")
write_rds(res4_df, "Output/DA_EC_P2_Grass.rds")
```
Heatmap of EC
Data transformation
```{r}
DA_EC_P2_Crop = read_rds("Output/DA_EC_P2_Crops.rds")
DA_EC_P2_Grass = read_rds("Output/DA_EC_P2_Grass.rds")
DA_EC_P2 = unique(c(DA_EC_P2_Crop$EC_number), c(DA_EC_P2_Grass$EC_number))
count_P2 = p2_EC_cleaned
count_P2 = microbiomeMarker::normalize(count_P2, method = "CLR")
count_P2 = count_P2 %>% as.data.frame() %>% filter(rownames(count_P2) %in% DA_EC_P2)
```
Distance between the variables (intergroups, Control vs Amended)
```{r}
dist_a_vs_c = vegan::vegdist(t(count_P2), method = "euclidean")
tree_a_vs_c = hclust(dist_a_vs_c, method = "complete")
cutree_a_vs_c = cutree(tree_a_vs_c, k = 5 )
table(cutree_a_vs_c)
```
Distance between the observation (EC)
```{r}
dist_ec = vegan::vegdist(count_P2, method = "euclidean")
tree_ec = hclust(dist_ec, method = "complete")
#cut into arbitrary clusters
cutree_ec = cutree(tree_ec, k = 8 )
table(cutree_ec)
```
Heatmap ##voir avec Domitille
```{r}
anno_col = as.data.frame(meta_cleaned[,c("groups","time")])
colnames(anno_col) = c("Groups","Time")
rownames(anno_col) = rownames(meta_cleaned)
anno_col$Groups = as.factor(anno_col$Groups)
anno_col$Time = as.factor(anno_col$Time)
anno_row = EC_number_class
anno_row = dplyr::select(anno_row, -EC_number)
time_colors <- c(
  "0" = "#1f77b4",   
  "3" = "#ff7f0e",   
  "51" = "#2ca02c",  
  "125" = "#d62728"  
)
ec_colors <- c("Hydrolases" = "#E41A1C",  # Red
  "Isomerases" = "#377EB8",  # Blue
  "Ligases" = "#4DAF4A",  # Green
  "Lyases" = "#984EA3",  # Purple
  "Oxidoreductases" = "#FF7F00",  # Orange
  "Transferases" = "#A65628")  # Yellow)
vect_colours = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd", 
                "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2")
list_colors = list(Groups = vect_colours, Time = time_colors, EC_class = ec_colors)
pheatmap(count_P2,
            cluster_rows = tree_ec, cluster_cols = tree_a_vs_c,
              show_rownames = FALSE, 
              cutree_rows = 8, #9, #5
              cutree_cols = 5,
              annotation_colors = list_colors,
              annotation_col = anno_col, 
              annotation_row = anno_row)
```
Part II : _Analysis of COG abundance in the output of P2_
tidy up the COG table
```{r}
p2_COG_cleaned = p2_COG[,rownames(meta_cleaned)]
```
PCA analysis for COG
```{r}
p2_COG_cleaned = round(p2_COG_cleaned,0)
dds_P2 = DESeqDataSetFromMatrix(countData = p2_COG_cleaned,
                                  colData = meta_cleaned,
                                  design = ~ groups + time, tidy = FALSE)
vst_P2 = vst(dds_P2,blind = FALSE) # variance stabilizing transformation
pcaData = plotPCA(vst_P2, intgroup = c("groups", "time"), returnData = TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(x = PC1, y = PC2, color = groups)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of COG abundance in output of PICRUSt2") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
ggsave("Output/PCA_COG_P2.png", width = 8, height = 10, dpi = 300)
```
DAG analysis about COG for Cropland
```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        COG_number = character())
p2_COG_cleaned = as.data.frame(p2_COG_cleaned)
for ( dataset in c("P2_crop_0","P2_crop_3" ,"P2_crop_51", "P2_crop_125")){              # Erratum : manque de la variable P2_crop_3
  counts_matrix <- as.matrix(p2_COG_cleaned[,get(dataset)],rownames = TRUE)
  sample_data <- meta_cleaned[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ modality, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$COG_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, COG_anno, by = "COG_number")
write_rds(res4_df, "Output/DA_COG_P2_Crops.rds")
```
DAG analysis about COG for Grassland
```{r}
res4_df <- data.frame( baseMean=numeric(), 
                        log2FoldChange=numeric(), 
                        lfcSE=numeric(), 
                        stat=numeric(), 
                        pvalue=numeric(), 
                        padj=numeric(),
                        conditions = factor(),
                        COG_number = character())
p2_COG_cleaned = as.data.frame(p2_COG_cleaned)
for ( dataset in c("P2_grass_0", "P2_grass_3", "P2_grass_51","P2_grass_125")){              # Erratum : manque de la variable P2_grass_125
  counts_matrix <- as.matrix(p2_COG_cleaned[,get(dataset)],rownames = TRUE)
  sample_data <- meta_cleaned[get(dataset),,drop=FALSE]
  res <- DESeqDataSetFromMatrix(countData = counts_matrix,
                                  colData = sample_data,
                                  design = ~ modality, tidy = FALSE)
 res <- DESeq(res, parallel = TRUE)
 res_1 = results(res, parallel = TRUE)
  res_2 = as.data.frame((res_1))
  res_2$COG_number = rownames(res_2)
  res_2$conditions = as.factor(rep(dataset, nrow(res_2)))
  res4_df <- bind_rows(res4_df, res_2)
}
rownames(res4_df) = NULL
res4_df = filter(res4_df, padj < 0.01 & abs(log2FoldChange) > 0.25)
res4_df = res4_df[order(res4_df$log2FoldChange, decreasing = TRUE),]
res4_df = left_join(res4_df, COG_anno, by = "COG_number")
write_rds(res4_df, "Output/DA_COG_P2_Grass.rds")
```

```{r}

```