---
title: "Test PCA et PCoA avec le jeu des données après filtré des ECs en commun"
auteur: "Dinh-Phuc Hoang"
date: "2025-04-03"
---


```{r}
library(tidyverse)
library(dplyr)
library(phyloseq)
library(DESeq2)
library(BiocParallel)
library(FactoMineR)
library(Hotelling)
library(microbiomeMarker)
register(SnowParam(4))
```

```{r}
tb_ec = read_tsv("Data FONCTIOMIC/counts_by_EC_number.norm.tsv")
meta = read_tsv("Data FONCTIOMIC/complete_metadata_table.tsv")
meta = as.data.frame(meta)
row.names(meta) = meta$sample_id
p2_EC = read_rds("Data FONCTIOMIC/picrust2out_ver2/p2_EC.rds")
p2_EC = p2_EC[,str_sort(colnames(p2_EC),decreasing = F)]
```

```{r}
meta$Category = as.factor(meta$Category)
meta$Category = relevel(meta$Category, ref = "Control Cropland")
meta$time = as.factor(meta$time)
meta$time = relevel(meta$time, ref = "0")
meta$wheat_straw_amendment = as.factor(meta$wheat_straw_amendment)
meta$wheat_straw_amendment = relevel(meta$wheat_straw_amendment, ref = "Control")
p2_EC_cleaned = p2_EC[,rownames(meta)]
tb_ec = as.data.frame(tb_ec)
row.names(tb_ec) = paste0("EC:", tb_ec$EC_number)
tb_ec_cleaned = tb_ec %>% filter(!str_detect(rownames(tb_ec), "-$"))
tb_ec_cleaned = tb_ec_cleaned[,-c(1,50)]
rm(tb_ec,p2_EC)
```

```{r}
overlapping_ec = intersect(rownames(tb_ec_cleaned), rownames(p2_EC_cleaned))
write_rds(overlapping_ec, "Data FONCTIOMIC/overlapping_ec.rds")
tb_ec_cleaned = tb_ec_cleaned[overlapping_ec,]
p2_EC_cleaned = p2_EC_cleaned[overlapping_ec,]
```
PCA analysis (VST transformation of DESeq2) of ECs in the output of MGS and PICRUSt2
- _MGS_
```{r}
tb_ec_vst = round(tb_ec_cleaned,0)
dds_MGS = DESeqDataSetFromMatrix(countData = tb_ec_vst,
                                  colData = meta,
                                  design = ~ Category + time, tidy = FALSE)
vst_MGS = vst(dds_MGS,blind = FALSE) # variance stabilizing transformation
#PCA plot
pcaData = plotPCA(vst_MGS, intgroup = c("Category", "time"), returnData = TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of MGS (VST)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))

```
- _P2_
```{r}
p2_ec_vst = round(p2_EC_cleaned,0)
dds_P2 = DESeqDataSetFromMatrix(countData = p2_ec_vst,
                                  colData = meta,
                                  design = ~ Category + time, tidy = FALSE)
vst_P2 = vst(dds_P2,blind = FALSE) # variance stabilizing transformation
#PCA plot
pcaData = plotPCA(vst_P2, intgroup = c("Category", "time"), returnData = TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of P2 (VST)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
```
PCA analysis (Data raw) of ECs in the output of MGS and PICRUSt2

```{r}
res.pca = PCA(tb_ec_cleaned, scale.unit = FALSE,graph = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(res.pca$eig[1:2,2], 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of MGS (Data raw)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
```

```{r}
res.pca = PCA(p2_EC_cleaned, scale.unit = FALSE,graph = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(res.pca$eig[1:2,2], 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of P2 (Data raw)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))

```
PCA analysis (CLR transformation) of ECs in the output of MGS and PICRUSt2

```{r}
tb_ec_clr = microbiomeMarker::normalize(tb_ec_cleaned, method = "CLR")
res.pca = PCA(tb_ec_clr, scale.unit = FALSE, graph = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(res.pca$eig[1:2,2], 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of MGS (CLR transfomation)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
```

```{r}
p2_ec_clr = microbiomeMarker::normalize(p2_EC_cleaned, method = "CLR")
res.pca = PCA(p2_ec_clr, scale.unit = FALSE, graph = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(res.pca$eig[1:2,2], 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of P2 (CLR transfomation)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))

```
PCA analysis (CLR transformation + Imputation) of ECs in the output of MGS and PICRUSt2

```{r}
tb_ec_clr1 = tb_ec_cleaned
tb_ec_clr1[tb_ec_clr1 == 0] = 0.01
tb_ec_clr1 = clr(tb_ec_clr1)
res.pca = PCA(tb_ec_clr1, scale.unit = FALSE,graph = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(res.pca$eig[1:2,2], 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of MGS (CLR transfomation + Imp)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
```

```{r}
p2_ec_clr1 = p2_EC_cleaned
p2_ec_clr1[p2_ec_clr1 == 0] = 0.01
p2_ec_clr1 = clr(p2_ec_clr1)
res.pca = PCA(p2_ec_clr1, scale.unit = FALSE,graph = FALSE)
pcaData = as.data.frame(res.pca$var$coord[,1:2])
pcaData$sample = rownames(pcaData)
pcaData = left_join(pcaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
colnames(pcaData)[1:2] = c("PC1","PC2")
percentVar <- round(res.pca$eig[1:2,2], 1)
ggplot(pcaData, aes(x = PC1, y = PC2, color = Category)) +
  geom_point(size =7) +
  geom_text(aes(label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
 ggtitle("PCA of EC abundance in output of P2 (CLR transfomation + Imp)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
```
PCoA analysis (Bray-Curtis distance) of ECs in the output of MGS and PICRUSt2
Raw data

```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(tb_ec_cleaned), method = "bray"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of MGS (Data raw)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
```

```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(p2_EC_cleaned), method = "bray"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of P2 (Data raw)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))

```

PCoA analysis (VST) of ECs in the output of MGS and PICRUSt2
```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(assay(vst_MGS)), method = "euclidean"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of MGS (VST)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
```

```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(assay(vst_P2)), method = "euclidean"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of P2 (VST)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))

```
PCoA analysis (CLR transformation) of ECs in the output of MGS and PICRUSt2
```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(tb_ec_clr), method = "euclidean"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of MGS (CLR transformation)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
```

```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(p2_ec_clr), method = "euclidean"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of P2 (CLR transformation)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
```
PCoA analysis (CLR transformation + imp) of ECs in the output of MGS and PICRUSt2
```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(tb_ec_clr1), method = "euclidean"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of MGS (CLR transformation + IMP)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
```

```{r}
res.pcoa = ape::pcoa(vegan::vegdist(t(p2_ec_clr1), method = "euclidean"))
pcoaData = as.data.frame(res.pcoa$vectors[,1:2])
pcoaData$sample = rownames(pcoaData)
pcoaData = left_join(pcoaData, meta[,c("Category","time","sample_id")], join_by(sample == sample_id), )
pcoaData.centroids = pcoaData %>% group_by(Category, time) %>% dplyr::summarise(mean.Axis.1 = mean(Axis.1), mean.Axis.2 = mean(Axis.2))
pcoaData = left_join(pcoaData, pcoaData.centroids, by = c("Category", "time"))
ggplot(pcoaData) +
  geom_point(aes(x=Axis.1, y=Axis.2, colour=Category), size=1) +
  geom_segment(aes(x=mean.Axis.1, y=mean.Axis.2, xend=Axis.1, yend=Axis.2, colour = Category))+
  geom_point(aes(x=mean.Axis.1, y=mean.Axis.2, colour=Category), size=6) +
  geom_text(aes(x=mean.Axis.1, y=mean.Axis.2,label = time), fontface = "bold", colour = "black", size = 3, alpha = 0.6)+
 ggtitle("PCoaA of EC abundance in output of P2 (CLR transformation + IMP)") +
    scale_colour_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                        guide=FALSE) + 
    scale_fill_manual(values = c("Amended Cropland"="#fbbb3f", "Control Cropland"="#f2e0bd",
                                  "Amended Grassland"="#2abf2a", "Control Grassland"="#b2d9b2"),
                      labels = c("Amended Cropland"='Amended Crops', "Control Cropland"='Control Crops', 
                                  "Amended Grassland"='Amended Grassland', "Control Grassland"='Control Grassland'))+
  guides(color = guide_legend(title = "Groups")) +  
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))
```